\documentclass[twoside, a4paper, 10pt]{amsart}
\title[ ]{Galois Theory notes}
%\usepackage{amsaddr}
%\email{Kamil.Bulinski@minetec.com.au}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{verbatim}
\usepackage{amsmath, amssymb}
\usepackage{tikz}
\usetikzlibrary{matrix, arrows}
\usepackage{listings}
\usepackage{color}
\usepackage{listings}
\usepackage[all]{xy}
\usepackage[pdftex,colorlinks,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{graphicx}
\usepackage{float}
\usepackage[margin=3cm]{geometry}
\usepackage{bigints}
\usepackage{dsfont}
\setlength{\textwidth}{6.5in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}
\linespread{1.3}


\setcounter{secnumdepth}{4}

\begin{document}
\maketitle
\raggedbottom


%% Mathcal large
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cK}{\mathcal{K}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}}
\newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\cU}{\mathcal{U}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cZ}{\mathcal{Z}}
%% Mathbb large
\newcommand{\bA}{\mathbb{A}}
\newcommand{\bB}{\mathbb{B}}
\newcommand{\bC}{\mathbb{C}}
\newcommand{\bD}{\mathbb{D}}
\newcommand{\bE}{\mathbb{E}}
\newcommand{\bF}{\mathbb{F}}
\newcommand{\bG}{\mathbb{G}}
\newcommand{\bH}{\mathbb{H}}
\newcommand{\bI}{\mathbb{I}}
\newcommand{\bJ}{\mathbb{J}}
\newcommand{\bK}{\mathbb{K}}
\newcommand{\bL}{\mathbb{L}}
\newcommand{\bM}{\mathbb{M}}
\newcommand{\bN}{\mathbb{N}}
\newcommand{\bO}{\mathbb{O}}
\newcommand{\bP}{\mathbb{P}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bS}{\mathbb{S}}
\newcommand{\bT}{\mathbb{T}}
\newcommand{\bU}{\mathbb{U}}
\newcommand{\bV}{\mathbb{V}}
\newcommand{\bW}{\mathbb{W}}
\newcommand{\bX}{\mathbb{X}}
\newcommand{\bY}{\mathbb{Y}}
\newcommand{\bZ}{\mathbb{Z}}


\newcounter{dummy} \numberwithin{dummy}{section}

\theoremstyle{definition}
\newtheorem{mydef}[dummy]{Definition}
\newtheorem{prop}[dummy]{Proposition}
\newtheorem{corol}[dummy]{Corollary}
\newtheorem{thm}[dummy]{Theorem}
\newtheorem{lemma}[dummy]{Lemma}
\newtheorem{eg}[dummy]{Example}
\newtheorem{notation}[dummy]{Notation}
\newtheorem{remark}[dummy]{Remark}
\newtheorem{claim}[dummy]{Claim}
\newtheorem{Exercise}[dummy]{Exercise}
\newtheorem{question}[dummy]{Question}
\newtheorem{conjecture}[dummy]{Conjecture}

\section{Splitting fields and Normal extensions}

\begin{prop} Let $K \leq L$ be fields. Suppose that $\alpha \in L$ is algebraic over $K$ and let $p(x) \in K[x]$ be a minimal polynomial for $\alpha$. Then there is a unique isomorphism $K[x]/(p(x)) \to K[\alpha] = K(\alpha)$ mapping $x$ to $\alpha$ and fixing $K$. \end{prop}

\begin{proof} There is a unique map $K[x] \to K[\alpha]$ mapping $x$ to $\alpha$ and fixing $K$. It is surjective and its kernel is the ideal generated by $p(x)$. \end{proof}

If $\sigma: K \to L$ is a homomorphism of fields and $f = a_0 + a_1x + \cdots + a_nx^n \in F[x]$, then we let $f^{\sigma} = \sigma(a_0) + \sigma(a_1) x + \cdots + \sigma(a_n)x^n \in F[x]$.

\begin{lemma} Suppose that $\sigma: K \to L$ is an isomorphism of fields and suppose $K' = K[\alpha]$ is an extension of $K$ where $f \in F[x]$ the minimal polynomial of $\alpha \in K'$. Let $\sigma:K \to L$ be a field homomorphism.

\begin{enumerate}
	\item If $\sigma': K' \to L$ extends $\sigma$, then $f^{\sigma}(\sigma'(\alpha)) = 0$
	\item If $\beta \in L$ satisfies that $f^{\sigma}(\beta) = 0$, then there is precisely one extension of $\sigma$ mapping $\alpha$ to $\beta$.
\end{enumerate}
\end{lemma}

\begin{proof} The first point is obvious. For the second point, let $\phi:K[x] \to L$ be given by $\phi(P) = P^{\sigma}(\beta)$. This is a ring homomorphism. Now observe that $\phi(f) = f^{\sigma}(\beta) = 0$, thus $\phi$ vanishes on the ideal generated by $f$ and so there is a well defined field homomorpism $\phi:K[x]/(f) \to L$ mapping $x + (f)$ to $\beta$. Finally, we use the isomorphism $ K' \cong K[x]/(f)$ that maps $\alpha$ to $x$ and fixes $K$, giving the desired extension. The extension is clearly unique as $K' = K(\alpha)$.

\end{proof}

\begin{prop} \label{prop: extend morphism to K'} Let $K \leq K'$ be an algebraic field extension and suppose that $\sigma:K \to L$ is a field homomorphism where $L$ is algebraically closed. Then there exists an extension $\sigma':K' \to L$. Moreover, $\sigma'$ must be an isomorphism if $K'$ is algebraically closed and $L$ is algebraic over $\sigma(K)$.

\end{prop}

\begin{proof} Use Zorn's lemma to construct a maximal subfield $K'' \subset K$ such that $\sigma$ extends to $K''$. If $K'' \neq K'$ then choose $\alpha \in K' \setminus K''$. Now as $K'$ ia algebraic over $K$ we can let $f \in K[x]$ be a minimal polynomial of $\alpha$ over $K$. Now as $f^{\sigma}$ has a root in $L$ as $L$ is algebraically closed, we can use the previous lemma to extend $\sigma$ to $K''[\alpha]$, contradicting the maximality of $K''$. If $K'$ is algebriacally closed, then so is $\sigma'(K')$ since any element of $\sigma'(K')[x]$ is of the form $f^{\sigma'}$ for some $f \in K'[x]$ and so we can let $\alpha$ be a root of $f$, giving that $\sigma'(\alpha)$ is a root of $f^{\sigma'}$. Now $\sigma'(K') \geq \sigma(K)$ so if $L$ is algebraical over $\sigma(K)$, then $L$ is also algebraic over $\sigma'(K')$. So if $L$ is algebraically closed then $L = \sigma'(K')$, giving that $\sigma'$ is surjective and thus an isomorphism (all field isomorphisms are injective). \end{proof}

\begin{corol} The algebraic closure of a field $K$ is unique upto an isomorphism fixing $K$.

\end{corol}

\begin{mydef}[Splitting field] Let $K \leq L$ be fields and let $\mathcal{F} \subset K[x]$ be a family of polynomials. We say that $L$ is a splitting field for $\mathcal{F}$ over $F$ if each $f \in \mathcal{F}$ splits into linear factors in $L[x]$ and $L$ is the field generated by $K$ and the roots of all polynomials in $\mathcal{F}$.

\end{mydef}

\begin{prop} A splitting field is unique upto an isomorphism fixing $F$.

\end{prop}

\begin{proof} Let $L \geq K$ and $L' \geq K$ be two splitting fields for a family $\mathcal{F} \subset K[x]$. We note that $L'$ and $L$ are both algebraic over $K$ (as they are generated by roots). This means that we may use Proposition~\ref{prop: extend morphism to K'} to extend the identity map $K \to K$ to a field homomorphism $\sigma:L \to \widehat{L'}$ where $\widehat{L'} \geq L'$ is algebraically closed. However, note that $\sigma(L) \subset L'$ since $\sigma$ maps each root of some $f \in \mathcal{F}$ to a root of $f$ (as $\sigma$ fixes $K$). So $\sigma:L \to L'$ is a homomorphism. It remains to show that $\sigma$ is surjective. To see this, let $f \in \mathcal{F}$ and write $f(x) = \prod_i (x - \alpha_i)$ where $\alpha_i \in L$. Then $f = f^{\sigma} = \prod_i(x - \sigma(\alpha_i))$. This shows that any root in $L'$ of any $f \in \mathcal{F}$ is in the image of $\sigma$ (using the unique factorization property). Thus as $L'$ is generated by these roots, the surjectivity of $\sigma$ follows. \end{proof}

If $K_1$ and $K_2$ are two fields with a common subfield $K$, we say that a homomorphism $K_1 \to K_2$ is a $K$-homomorphism if it restricts to the identity on $K$.

\begin{thm} Let $L$ be an algebraic extension of a field $K$. Then the following are equivalent.

\begin{enumerate}
	\item $L$ is a splitting field for some family of polynomials in $K[x]$.
	\item Any $K$-homomorphism $L \to \overline{L}$, where $\overline{L} \geq L$ is an algebriac closure, restricts to an automorphism of $L$
	\item Any irreducible polynomial in $K[x]$ that has a root in $L$ must decompose into linear factors in $L[x]$.
\end{enumerate}

\end{thm}

\begin{proof} (i) $\implies$ (ii): If $L$ is a splitting field for some polynomials in $K[x]$ and $\sigma:L \to \overline{L}$ is a $K$-homomorphism, then as in the proof of the uniqueness of splitting fields above, we see that $\sigma$ maps into $L$. We also saw that it permutes the roots of a polynomial in $K[x]$ in $L$ and thus the image of $\sigma$ is $L$, thus $\sigma$ is surjective and hence an automorphism.

(ii) $\implies$ (iii): Suppose $f \in K[x]$ is irreducible and has a root $\alpha \in L$. Now if $\alpha' \in \overline{L}$ is another root of $f$, then since $f$ is irreducible we have an isomorphism $ K[\alpha] \to K[\alpha']$ mapping $\alpha$ to $\alpha'$, which we may extend to an $K$-homomorphism $\sigma:L \mapsto \overline{L}$ by a previous Lemma. By condition $(ii)$, we see that $\sigma$ maps $L$ to $L$ and thus $\alpha' = \sigma(\alpha) \in L$. Hence $L$ contains all the roots of $f$.

(iii) $\implies$ (ii): As $L$ is algebraic, every element $\alpha \in L$ is the root of some irreducible polynomial $f \in K[x]$. We thus let $\mathcal{F} \subset K[x]$ be those irreducible polynomials with at least one root in $L$, which split into linear factors by assumption. Thus $L$ is the splitting field of $\mathcal{F}$ over $K$.

\end{proof}

\begin{mydef} We say that an extension $K \leq L$ is normal if it is the splitting field of some family of polynomials.

\end{mydef}

\begin{eg} The extension $\bQ \leq \bQ[2^{1/3}]$ is not normal. To see this we use the characterization (iii) in the Theorem as follows: The polynomial $x^3 - 2$ is irreducible, has one root $2^{1/3}$ in our extension but not any other. Alternatively, we can use (ii) by noting that although there is $\bQ$-homomorphism $\bQ[2^{1/3}] \to \overline{\bQ}$ mapping $2^{1/3}$ to $2^{1/3} e^{2\pi i /3}$, it does not restrict to an automorphism of $\bQ[2^{1/3}]$.

\end{eg}

\begin{eg} Normal is not transitive. As an example, consider the field extensions $\bQ \leq \bQ[\sqrt{2}] \leq \bQ[2^{1/4}]$. The intermediate field extensions are normal (as they are of degree $2$) but the extension $\bQ \leq \bQ[2^{1/4}]$ is not.

\end{eg}

\begin{mydef} If $L \geq K$ is an algebraic extension, then we say that $L' \leq L \leq K$ is a normal closure of $L \geq K$ if $L' \geq K$ is a normal extension and any $L' \geq L'' \geq K$ such that $L'' \geq K$ is normal must satisfy $L'' = L'$. That is, the normal closure if a minimal normal extension.

\end{mydef}

\begin{prop} \label{prop: normal closure existence} Every algebraic extension $L \geq K$ has a normal closure. More precisely, let $\mathcal{F}$ be the set of all irreducible polynomials in $K[x]$ such that each element of $L \setminus K$ is the root of some $f \in \mathcal{F}$. Then the splitting field of $\mathcal{F}$ is the normal closure of $L \geq K$.

\end{prop}

\begin{proof} Let $\overline{L} \geq L$ be the algebraic closure of $L$. Define $\overline{L} \geq L' \geq L$ to be the splitting field for the family $\mathcal{F} \subset K[x]$ of minimal polynomials for elements of $L$. We claim that $L'$ is the normal closure. Thus suppose that $L \leq L'' \leq L'$ is such that $K \leq L''$ is normal. We must show that $L'' = L'$, and since $L'$ is generated by the roots of elements of $\mathcal{F}$, we must show that any root $\alpha \in L'$ of a polynomial $f \in \mathcal{F}$ is in $L''$. To see this, note that by definition $f$ is a minimal polynomial of some $\alpha' \in L$. There is a $K$-homomorphism $\sigma:K[\alpha'] \to \overline{L}$ mapping $\alpha'$ to $\alpha \in L$. As $L'' \geq L \geq K[\alpha']$, we may extend this $K$-homomorphism to $\sigma:L'' \to \overline{L}$. But by characterization (ii) of the normality of $K \leq L''$, we see that $\sigma$ is an automorphism of $L''$. This means that $\alpha = \sigma(\alpha') \in L''$ as $\alpha' \in L \subset L''$. Thus this shows that $L' \subset L''$, and so $L' = L''$ as required. \end{proof}

\begin{prop} If $K \leq L$ is an algebraic extension and $L \leq L_1, L_2 \leq \overline{L}$ are two normal extensions of $K$, then $L_1 \cap L_2$ is a normal extension of $K$. In particular, if $L_1$ and $L_2$ are both normal closures of $L \geq K$, then $L_1 = L_2$.

\end{prop}

\begin{proof} This follows from characterization (iii): If $f \in K[x]$ is irreducible and has a root in $\alpha \in L_1 \cap L_2$, then $f$ decomposes to linear factors in $L_i[x]$ for $i=1,2$. By uniqueness of factorizations, this means that these linear factors are in $(L_1 \cap L_2)[x]$. \end{proof}

\begin{prop} A normal closure of an algebraic extension $L \geq K$ is unique upto an $L$-automorphism.

\end{prop}

\begin{proof} By the previous construction, we have one such normal closure given by $L[\mathcal{R}]$ where $$\mathcal{R} = \{ r \in \overline{L} ~|~ f(r) = 0 \text{ for some } f \in \mathcal{F} \}$$ where $\mathcal{F} \subset K[x]$ is the set of all irreducible polynomials such that each element of $L$ is the root of some $f \in \mathcal{F}$. We now let $L' \geq L$ be another field such that $L' \geq K$ is the normal closure of $L \geq K$. We now construct an isomorphism $L[\mathcal{R}] \to L'$ which fixes $L$. We extend the inclusion $L \to \overline{L'}$ to an $L$-homomorphism $\sigma:L[\mathcal{R}] \to \overline{L'}$. Note that $L'' = \sigma(L[\mathcal{R}]) = L[\sigma(\mathcal{R})]$ contains $L$ and is the splitting field of $\mathcal{F}$ in $\overline{L'}$ over $K$. Thus $L'$ and $L''$ are subfields of $\overline{L}$ that are normal extensions of $K$ and both contain $L$. Moreover, $L''$ is also a normal closure of $L \geq K$ as it follows the construction given in Proposition~\ref{prop: normal closure existence} (i.e., it is a splitting field of minimal polynomials over $K[x]$ of elements in $L$). By the previous proposition, it follows that $L' = L''$, thus $\sigma$ is an isomorphism. \end{proof}

\section{Seperable extensions}

\begin{lemma} An irreducible polynomial $f \in K[x]$ splits into distinct linear factors in some algebraic closure if and only if $f' = 0$.

\end{lemma}

\begin{proof} By the product rule it follows that if $f(\alpha) = 0$ then $\alpha$ is a repeated root if and only if $f'(\alpha) = 0$. If $f$ is irreducible, has a repeated root $\alpha$ and $f' \neq 0$ then $(X-\alpha) | gcd(f, f') | f$, which contradicts the irreducibility of $f$. \end{proof}

As a consequence, if $char K = 0$ then an irreducible polynomial must split into distinct linear factors.

\begin{mydef} We say that $f \in K[x]$ is seperable if $f$ splits into distinct linear factors in some (hence any) algebraic closure of $K$.

\end{mydef}

\begin{thm} If $char K = p$ and $f \in K[x]$ is irreducible, then each root of $f$ has multiplicity $p^r$ where $r$ is minimal non-negative integer such that $f(x) = g(x^{p^r})$ for some $g \in K[x]$.

\end{thm}

\begin{proof} Write $g(x) = \sum_j c_j x^j$. Since $$g'(x) = \sum_{j} j c_j x^j$$ we observe that $g'(x)$ is not the zero polynomial as follows: If $g'(x) = 0$ then $c_j = 0$ whenever $j$ is not divisible by $p$. From this it follows that $g(x) = \sum_k c_{kp} x^{kp} = h(x^p)$. It now follows that $$f(x) = g(x^{p^r}) = h((x^{p^r})^p) =  h(x^{p^{r+1}}),$$ which contradicts the maximality of $r$. Thus $g'(x) \neq 0$. This means that $g(x) = \prod_i (x - \alpha_i)$ where $\alpha_i$ are distinct. Write $\alpha_i = \beta_i^{p^r}$, which exists in an algebraic closure. Note that the $\beta_i$ must also be distinct. Thus $$f(x) = \prod_i (x^{p^r} - \beta_i^{p^r}) = \prod_i (x- \beta_i)^{p^r},  $$ where the last equality follows from Freshman's dream in characteristic $p$. As the $\beta_i$ are distinct, the proof is complete. \end{proof}

\begin{mydef} If $K \leq L$ is an algebraic field extension then $\alpha \in L$ is called seperable over $K$ if the minimal polynomial is seperable (splits over linear factors in some, hence any, algebraic closure). We say that the extension $K \leq L$ is seperable if all elements of $L$ are separable over $K$.

\end{mydef}

Thus from above, in characteristic zero all algebraic extensions are seperable, as all irreducible polynomials are seperable.

\begin{eg} Consider the field $K = \bF_p(t)$. The polynomial $f(x) = x^p - t$ is irreducible by Eisenstein's criterion in $\bF_p[t]$ as $t$ is prime in this UFD, and hence $f(x)$ is irreducible also over its field of fractions $K$ by Gauss's Lemma. Now, $f(\alpha) = 0$ in for some $\alpha \in \overline{K}$, that is $\alpha^p = t$. But by Freshman's dream we have that $$x^p - t = x^p - \alpha^p = (x-\alpha)^p,$$ thus $\alpha$ is a root of multiplicity $p$ for $f(x)$. Thus $f(x)$ is irreducible but not seperable. \end{eg}

\begin{mydef} If $K \leq L$ is an algebraic extension, then we let $$Hom_K(L, \overline{K})$$ denote the set of all $K$-homomorphisms $L \to \overline{K}$. We let $$|L:K|_s = |Hom_K(L, \overline{K})|$$ be the seperable degree of $K \leq L$, which does not depend on the choice of $\overline{K}$.

\end{mydef}

\begin{prop} \label{prop: Hom tower bijections} If $K \leq L \leq M$ are algebraic extensions then there is a bijection $$Hom_K(L, \overline{K}) \times Hom_L(M, \overline{K}) \to Hom_K(M, \overline{K}).$$ In particular $$|M:K|_s = |L:K|_s |M:L|_s.$$

\end{prop}

\begin{proof} For each $\sigma \in Hom_K(L, \overline{K})$ we choose an arbitrary (there are many choices) $\phi(\sigma):\overline{K} \to \overline{K}$ automorphism that extends $\sigma$, where we have used Proposition???. Now we define a mapping $$Hom_K(L, \overline{K}) \times Hom_L(M, \overline{K}) \to Hom_K(M, \overline{K})$$ by $$(\sigma, \tau) \mapsto \phi(\sigma) \circ \tau.$$ Let us first check that it is well defined. If $k \in K$ then $$(\phi(\sigma) \circ \tau) (k) = \phi(\sigma) (\tau(k)) = \phi(\sigma) (k) = \sigma(k) = k,$$ so indeed $\phi(\sigma) \circ \tau$ is a $K$-homomorphism. To show injectivity, suppose that $$\phi(\sigma) \circ \tau = \phi(\sigma') \circ \tau'.$$ Then for any $\ell \in L$ we have that $$\phi(\sigma)(\tau(\ell)) = \phi(\sigma)(\ell) = \sigma(\ell)$$ and by the same arugment $\phi(\sigma')(\tau'(\ell)) = \sigma'(\ell)$. Thus $\sigma = \sigma'$. This means that $\phi(\sigma) = \phi(\sigma')$ and so by injectivity of field automorphisms, we must have that $\tau'(m) = \tau(m)$ for all $m \in M$. So $\tau = \tau'$. It now remains to show injectivity. Thus suppose that $\gamma \in Hom_K(M, \overline{K})$. Let $\sigma$ be the restriction of $\gamma$ to $L$ and observe that $\sigma \in Hom_K(L, \overline{K})$. Now let $$\tau = \phi(\sigma)^{-1} \circ \gamma:M \to \overline{K}.$$ If $\ell \in L$ then $$\tau(\ell) = \phi(\sigma)^{-1}(\gamma(\ell)) = \phi(\sigma)^{-1}(\sigma(\ell)) = \phi(\sigma)^{-1} \phi(\sigma)(\ell) = \ell,$$ thus indeed $\tau \in Hom_L(M, \overline{K})$. This shows that $\gamma = \phi(\sigma) \circ \tau$ is in the image of our map, thus our map is surjective. \end{proof}

\begin{prop}\label{prop: seperable degree prime power} If $K \leq L$ is a finite extension then 
\begin{enumerate}
	\item If $K$ has characteristic zero then $|L:K| = |L:K|_s$
	\item If $K$ has characteristic $p$ then $|L:K| = p^r|L:K|_s$ for some integer $r \geq 0$.
\end{enumerate}
\end{prop}

\begin{proof} By finiteness of this extension $L$ can be obtained from $K$ by finitely many simple extensions, so we only need to prove this when $L = K(\alpha)$ is a simple extension and then use the previous proposition to give the general case by induction. If $Char K = 0$ then we know that $|L:K| = deg f = |L:K|_s$ where $f \in K[x]$ is the minimal polynomial of $\alpha$, where we have used the fact that $f$ is seperable and there is a unique $K$-homomorphism mapping $\alpha$ to any given root of $f$. If $Char K = p$ then $|L:K| = deg f = p^r |L:K|_s$ where $r$ is maximal integer such that $f(x) = g(x^{p^r})$ for some polynomial $g(x) \in K[x]$, as seen in a previously proven result. Thus completing the proof. \end{proof}

\begin{thm}\label{thm: characterise seperable} Let $K \geq L$ be a finite extension. The following are equivalent.

\begin{enumerate}
	\item $K \geq L$ is seperable.
	\item $L = K(a_1, \ldots, a_n)$ for some $a_1, \ldots, a_n \in L$ that are separable over $K$
	\item $|L:K|_s = |L:K|$
\end{enumerate}

\end{thm}

\begin{proof} (i) $\implies$ (ii) is trivial. (ii) $\implies$ (iii): Letting $K_i = K_{i-1}(a_i)$ we see that $a_i$ is separable over $K_{i-1} \geq K$and thus $|K_{i}:K_{i-1}| = deg(f_i) = |K_i:K_{i-1}|_s$ where $f_i$ is the minimal polynomial of $a_i$ over $K_{i-1}$. We are now done by the multiplicativity formula. (iii) $\implies$ (i): We only need to focus on $Char K = p >0$. If $a \in L$ is not seperable over $K$ then $$|K(a):K|_s < |K(a):K|$$, but then $$|L:K|_s = |L:K(a)|_s |K(a):L|_s < |L:K(a)||K(a):K| = |L:K|.$$ 

\end{proof}

\begin{corol} If $K \leq L \leq M$ are algebraic extensions then $K \leq M$ is seperable if and only if $K \leq L$ and $L \leq M$ are seperable.

\end{corol}

\begin{proof} First suppose $K \leq M$ is seperable. Then clearly $K \leq L$ is seperable. Now for $a \in M$ we have that the minimal polynomial $f(x) \in K[x]$ of $a$ over $K$ splits into linear factors. If $g(x) \in L[x]$ is the minimal polynomial of $a$ over $L$, then clearly $g(x) | f(x)$ as $f(x) \in L[x]$. Thus $g(x)$ also splits into linear factors.

Conversely, assume now that $K \leq L$ and $L \leq M$ are seperable. Fix $a \in M$. Then $|L(a):L| = |L(a):L|_s$ as $L \leq M$ is seperable. Now let $L' \leq L$ be the field generated by $K$ and the coefficients of the minimal polynomial $f(x) \in L[x]$ of $a$ over $L$. Thus $f(x) \in L'[x]$ which means that $a$ is seperable over $L'$ as well (as $f(x)$ splits into linear factors and $f(a) = 0$). Thus $|L'(a):L'|_s = |L(a):L|$. It now follows that $$|L'(a):K|_s = |L'(a):L'|_s |L':K|_s = |L'(a):L||L':K| = |L'(a):K|,$$ hence be the previous theorem we have that $L'(a)$ is seperable over $K$, and thus $a$ is seperable over $K$. \end{proof}

\begin{thm}[Primitive element theorem] If $K \leq L$ is a finite seperable extension, then $L = K(a)$ for some $a \in L$. \end{thm}

\begin{proof} If $L$ is finite, then this follows from the fact that the multiplicative group of a field is cyclic. Suppose thus that $K$ and $L$ are infinite. We may reduce to the case where $L = K(\alpha, \beta)$, as the general case then follows by induction (If $L = K(a_1, \ldots, a_n)$ then $L = K'(a_1,a_2)$ where $K' = K(a_3, \ldots, a_n)$ and certainly $L$ is seperable over $K'$). For $c \in K$, we let $\gamma_c = \alpha + c \beta$. We will show that $L = K(\gamma_c)$ for infinitely many $c \in K$ as follows. If $L \neq K(\gamma_c)$ then definitely $\beta \notin K(\gamma_c)$. As $L$ is seperable over $K(\gamma_c)$, this means that the minimal polynomial of $\beta$ over $K(\gamma_c)$ has another root $\beta' \in \overline{K}$. Thus there exists a $K(\gamma_c)$-homomorphism $\sigma:L \to \overline{K}$ with $\sigma(\beta) = \beta' \neq \beta$. We thus get that $$\sigma(\alpha) + c \sigma(\beta) = \alpha + c \beta$$ and thus $$c = \frac{\sigma(\alpha) - \alpha}{\beta - \sigma(\beta)}.$$ But the right hand side has only finitely many choices (as there are only finitely many choices of $\sigma$) and so if we choose a $c$ not of this form (as $K$ is infinite) we see that $L = K(\gamma_c)$ as desired. \end{proof}

\section{Galois Extensions}

\begin{mydef} A field extension $K \leq L$ is called \textit{Galois} if it is normal and seperable. We also say $L$ is \textit{Galois} over $K$. We define $Gal(L/K) := Aut_K(L)$ to be the set of $K$-automorphisms $L \to L$.

\end{mydef}

\begin{prop} Suppose that $K \leq L$ is Galois and $K \leq E \leq L$ is an intermediate field. 

\begin{enumerate}
	\item Then $L$ is also Galois over $E$ and $Gal(L/E) \subset Gal(L/K)$.
	\item If $E$ is also Galois over $K$, then every $\sigma \in Gal(L/K)$ restricts to an automorphism $\sigma|_E \in Gal(E/K)$. Moreover, this restriction homomorphism is surjective.
\end{enumerate}
\end{prop}

\begin{proof} It is clear from the definition that $K \leq L$ normal (resp. seperable) implies that $E \leq L$ is normal (resp. seperable). Any $E$-automorphism fixes each element of $e$ and hence each element of $K \leq E$, thus the inclusion in (i). For (ii): We already know that as $E$ is normal over $K$ then any $K$-automorphism of $L$ must map $E$ into $E$ surjectively as $E$ is the splitting field of some polynomials over $K$ and thus any automorphism permutes these roots (which are the generators for $E$ over $K$). By Proposition??? any $K$-automorphism $\sigma:E \to E$ can be extended to some $K$-homomorphism $\overline{\sigma}:\overline{K} \to \overline{K}$. But $\overline{\sigma}$ must permute the roots of any polynomial in $K[x]$ and in particular those for which $L$ is the splitting field for, thus $\overline{\sigma}$ restricts to an automorphism of $L$. \end{proof}

\begin{prop}\label{prop: fixed field galois group} Let $L$ be a field and let $G$ be a subgroup of $Aut(L)$. Let $$K=L^G:= \{a \in L ~|~ ga = a \text{ for all } g \in G\}$$ be the fixed field of $G$. 
\begin{enumerate}
	\item If $G$ is finite then $K \leq L$ is a finite Galois extension and $Gal(L/K) = G$ and $|L:K| = |G|$
	\item If $K \leq L$ is algebraic and $G$ is not necessarily finite, then $K \leq L$ is a Galois Extension with $G \leq Gal(L/K)$. 
\end{enumerate}

\end{prop}

\begin{proof} We first show that in  both case (i) or (ii), the orbit $Ga$ is finite for all $a \in L$. This is obvious in (i). In (ii), since $a$ is algebraic over $K$ then there is a non-zero polynomial $f \in K[x]$ such that $f(a) = 0$. But now $f(g(a)) = 0$ for all $g \in G$ as $g$ fixes $K$ and hence $f$. Thus the orbit $Ga$ is contained in the roots of $f$, which is a finite set. So now we just assume that $Ga$ is finite for all $a \in L$. Consider the polynomial $$f_a(x) = \prod_{\alpha \in Ga} (x - \alpha).$$ Note that $g$ permutes these linear factors, thus $f_a(x) \in L^G[x] = K[x]$. Thus $a$ is algebraic over $K$. Moreover, it now follows that $L$ is the splitting field of $\{f_a ~|~ a \in L\}$, thus $L$ is normal over $K$ and also seperable as these factors are distinct. Thus $K \leq L$ is indeed a Galois extension.  We now complete the proof of (i), thus assume from now that $G$ is finite. To show that $K \leq L$ is a finite extension, it will be enough to find a uniform bound on intermediate fields $K \leq L' \leq L$ such that $K \leq L'$ is a finite normal extension (because we know $K \leq L$ is algebraic and thus if it is infinite then we choose finitely many elements in $L$ such that the field they generate is arbitrarily large. The normal closure of this field is also finitely generated hence a finite extension). Now as such an $L'$ is finite, the primitive root theorem says that $L' = K(a)$ for some $a \in L$. But then we know that the minimal polynomial of $a$ is a divisor of $f_a(x) \in K[x]$ above, which is of degree at most $|G|$, thus $|L':K| \leq |G|$. It follows that $|L:K| \leq |G|$, so $L$ is indeed a finite extension. Now we use the primitive root theorem to write $L = K(\alpha)$ for some $\alpha \in L$. Observe that if $g \alpha = \alpha$ then $g = Id_L = 1_G$, thus $|G| \leq |L:K|_s = |L:K|$. This completes the proof that $|L:K| = |G|$. \end{proof}

\begin{thm}[Fundamental theorem of Galois Theory] Suppose that $K \leq L$ is a Galois extension. Let $Fields(L/K)$ denote the set of intermediate fields $K \leq E \leq L$. For a group $G$ we let $SubGrps(G)$ denote the set of subgroups $H \leq G$. Define the maps $$\phi: SubGrps(Gal(L/K)) \to Fields(L/K)$$ that maps $$H \leq Gal(L/K)$$ to the fixed field $L^H$ and $$\psi:Fields(L/K) \to SubGrps(Gal(L/K)) $$ which maps an intermiediate field $K \leq E \leq L$ to the Galois group $Gal(L/E) = Aut_E(L)$. Then $$\phi \circ \psi = Id_{Fields(L/K)}.$$ Moreover, if the extension $K \leq L$ is finite, then $$\psi \circ \phi = Id_{SubGrps(Gal(L/E))}$$ and thus these maps bijective and inverses of each other. Moreover, if $K \leq L$ is finite then a subgroup $H \leq Gal(L/K)$ is normal if and only if $L^H$ is normal over $K$ (and thus $K \leq L^H$ is Galois), in which case there is a surjective group homomorphism $Gal(L/K) \to Gal(L^H/K)$ which maps $\sigma$ to $\sigma|_{L^H}$ and $H$ is the kernel of this map, so $$Gal(L/K)/H \cong Gal(L^H/K).$$

\end{thm}

\begin{proof} Let $K \leq E \leq L$ be an intermediate field, then we know that $E \leq L$ is Galois. Now let $H = Gal(L/E)$ and $E' = L^H$. Clearly $E \leq E'$ (if $a \in E$ then $h(e) = e$ for all $h \in Gal(L/E)$ and so $e \in L^H = E'$). Now suppose for contradiction that $a \in E'$ but $a \notin E$. Hence as $L/E$ is seperable, the minimal polynomial of $a$ over $E$ has another root $b \neq a$ and thus there is a $h \in Aut_E(L) = H$ that maps $a$ to $b$. Thus $a \notin L^H = E'$, a contradiction. This means that $E' = E$, thus showing that $\psi \circ \phi$ is the identity as claimed.

Now we assume that $K \leq L$ is finite, thus $L = K(\alpha)$ for some $\alpha \in L$ by the primitive root theorem. Clearly $G = |Gal(L/K)|$ is finite since $g \in G$ is uniquely determined by the image of $\alpha$, which must be a root of the minimal polynomial of $\alpha$. Choose a subgroup $H \leq Gal(L/K)$. Thus $H$ is finite and we may apply the Proposition~\ref{prop: fixed field galois group} to deduce that $\psi(\phi(H)) = Gal(L/L^H) = H$. Thus $\phi$ and $\psi$ are inverses in when $K \leq L$ is a finite extension.

Finally, suppose that $K \leq E \leq L$ is such that $E$ is a normal extension of $K$. We now wish to show that $H = Gal(L/E)$ is normal in $Gal(L/K)$. To see this, we know from Propoistion ??? that there is a surjective homomorphism $Gal(L/K) \to Gal(E/K)$ mapping $\sigma \in Gal(L/K)$ to $\sigma|E$. Observe that $g \in Gal(L/K)$ is in the kernel of this homomorphism if and only if $g|E = 1$ which happens if and only if $g(e) = e$ for all $e \in E$ which happens if and only if $g \in Gal(L/E) = H$. Thus $H$ is the kernel of a homomorphism, thus a normal subgroup as desired.

Conversely, suppose that $H$ is a normal subgroup of $Gal(L/K)$ and let $E = L^H$. We wish to show that $L^H$ is normal over $K$. Thus we wish to show that if $\sigma:L^H \to \overline{K}$ is a $K$-homomorphism then $\sigma(L^H) = L^H$. To show this, let $a \in L^H$ be arbitrary and let $b = \sigma(a)$. To show $b \in L^H$ we have to show that $hb = b$ for all $h \in H$. Now extend $\sigma$ to an automorphism $\sigma:L \to L$ (as $L$ is normal over $K$). Then $\sigma H = H \sigma$ as $H$ is normal in $Gal(L/K)$. Thus $h \sigma = \sigma h'$ for some $h' \in H$ and thus $$ h b = h \sigma a = \sigma h'a = \sigma a = b.$$ Thus $b \in L^H$. So $\sigma(L^H) \subset L^H$. It now remains to show the opposite inclusion. Thus suppose $a \in L^H$, then $\sigma^{-1}H = H \sigma^{-1}$  (note that $\sigma^{-1}:L \to L$ is defined as $\sigma$ is an automorphism of $L$). Now the same argument shows that $\sigma^{-1}(a) \in L^H$ and thus $\sigma^{-1}(L^H) \subset L^H$, i.e., $L^H \subset \sigma(L^H)$. \end{proof}

\begin{eg} Let $\alpha = 2^{1/4}$ and let $L = \bQ[\alpha, i]$ which is the splitting field of the polynomial $X^4 - 2$. Let as compute the Galois group $G=Gal(L/\bQ)$. Obseve that for $g \in G$ we have that $$g(\alpha) \in \{\alpha, i \alpha, -\alpha, -i \alpha\}$$ and $$g(i) \in \{\pm i\}$$. Thus $|G| \leq 8$. Let us show that all $8$ combinations are possible (realised by some $g \in G$). Let $\sigma:L \to L$ be the complex conjugation map, so $\sigma \in G$. Now we know that for each $k \in \{0,1,2,3\}$ there exists a $g_k \in G$ such that $g(\alpha) = i^k \alpha$ (as $X^4-2$ is irreducible over $\bQ$ there is a $\bQ$-automorphism mapping any root to any other root). Now notice that $g_k \circ \sigma(\alpha) = g_k(\alpha) = i^k \alpha$ and yet $g_k \circ \sigma (i) = g_k(-i) = -g_k(i)$. Thus the elements $g_k \circ \sigma^{e} \in Gal(L/K)$ are all distrinct for distinct $(k,e) \in \{0,1, 2, 3\} \times \{0,1\}$ and so all $8$ combinations are possible. Let $ r \in G$ be the map given by $g(\alpha) = i\alpha$ and $g(i)=i$. Thus $g(i^k \alpha) = i^{k+1}\alpha$. So $r$ rotates the elements $\alpha, i\alpha, i^2\alpha, i^3\alpha$ cylically. While $\sigma$ is an involution that swaps $i\alpha$ with $i^3\alpha$ and fixes $\alpha, i^2\alpha$. Every element of $G$ as of the form $r^k \sigma^e$ where $(k,e) \in \{0,1,2,3\} \times \{0,1\}$. Thus $G$ is isomorphic to $D_8$ since if consider the elements $\alpha, i\alpha, i^2\alpha, i^3\alpha$ as succesive corners of a square, then $r$ is a rotation and $\sigma$ is a reflection. Note that $|L:\bQ| = 8$ and a $\bQ$-basis is given by $$\{\alpha^k i ^e ~|~ i \in \{0,1,2,3\}, e \in \{0,1\} \}.$$ 

Let us consider some intermediate fields and corresponding subgroups. First, consider the reflection group $\{1, \sigma\}$. The only elements of $L$ fixed by this group are $L \cap \bR = \bQ[2^{1/4}]$. This subgroup is not normal and indeed $\bQ[2^{1/4}]$ is not a normal extension of $\bQ$. On the other hand, the rotation sugroup $\langle r \rangle$ is normal, and so the fixed field should be normal. To compute the fixed field, note that we may write each $x \in L$ as $$x = \sum_{k=0}^3 \lambda_k \alpha^k,$$ for some unique $\lambda_k \in \bQ[i]$. Thus if $rx=x$ then $\lambda_k = i^k \lambda_k$, thus we must have that $x = \lambda_0 \in \bQ[i]$. This shows that the fixed field for this rotation subgroup is $\bQ[i]$, which indeed is normal (the splitting field of $x^2+1$). Observe now that each $g \in G$ restricts to an automorphism of this fixed field $\bQ[i]$ and it restricts to the identity on $\bQ[i]$ if and only if $g$ is in this rotation group, giving the isomorphism $$G/\langle r \rangle \cong Gal(\bQ[i]/\bQ).$$

Now consider the group of order $4$ generated by the reflections $\sigma$ (complex conjugation) and the map $\tau \in G$ given by $\tau(\alpha) = -\alpha$ and $\tau(i) = i$. The subgroup is abelian of order $4$ as $\sigma$ and $\tau$ commute. The fixed field is $\bQ[\alpha^2] = \bQ[\sqrt{2}]$, which is also normal over $\bQ$ (splitting field of $X^2-2$). \end{eg}

If $E,E' \leq L$ are two subfields, then we let $E \cdot E'$ denote the subfield of $L$ generated by these two subfields, i.e., that smallest subfield containing both. Explicitly, $$E \cdot E' = \{ \sum_{i=1}^n e_i e_i' ~|~ n \in \bZ_{>0} e_i \in E, e_i' \in E' \}.$$

\begin{corol} If $K \leq L$ is a finite Galois extension such that $K\leq E, E' \leq L$ are subfields and $H = Gal(L/E)$ and $H' = Gal(L/E')$ are the corresponding galois groups. Then we have that 
\begin{enumerate}
	\item $E \subset E'$ if and only if $H \subset H'$.
	\item $E \cdot E' = L^{H \cap H'}$
	\item $E \cap E' = L^{H''}$ where $H''$ is the smallest subgroup containing both $H$ and $H'$.  
\end{enumerate}
\end{corol}

\begin{proof} (i) is clear from the Galois correspondence. For (ii), note that if $e \in E$ and $e' \in E$ and $h \in H \cap H'$< then $h(ee') = h(e)h(e') = ee'$, thus $ee' \in L^{H \cap H'}$. Hence $E \cdot E' \subset L^{H \cap H'}$. For the reverse inclusion, note that if $h \in Gal(L / E \cdot E')$ then $h(e) = e$ and $h(e') = e'$ for all $e \in E, e' \in E'$ as $e,e' \in E \cdot E'$. Thus $h \in H \cap H'$. Thus part $(i)$ and the Galois correspondence shows that $E \cdot E' \supset L^{H \cap H'}$. For (iii), note that if $e \in E \cap E'$ then $h(e) = e$ and $h'(e) = e$ for all $h \in H$ and $h' \in H'$, thus $e$ is fixed by all products of elements in $H$ or $H'$, thus fixed by all elements in $H''$. This shows that $E \cap E' \subset L^{H''}$. Conversely, as $H \leq H''$, then $E = L^H \supset L^{H''}$. Likewise, $E' \supset L^{H''}$ as $H' \leq H''$. Thus $E \cap E' \supset L^{H''}$. \end{proof}

\begin{eg} Continuing from the previous examples, let $K = \bQ$, $L$ be the splitting field of $X^4-2$ and let $E = \bQ[2^{1/4}]$ and $E' = \bQ[i]$. We already saw the corresponding Galois groups are $H = \{1, \sigma\}$ and $H' = \{1,r,r^2,r^3\}$. Now $E \cdot E' = L$ but $H \cap H' = \{1\}$. The full Galois group is generated by these two groups and indeed $E \cap E' = \bQ$ is the corresponding Galois subgroup.\end{eg}

\begin{prop} Suppose $K \leq E, E'$ are finite Galois extensions where $E,E' \leq L$ for some field $L$. Then $E \cdot E'$ is also a finite Galois extension over $K$. Also:
\begin{enumerate}
	\item The restriction map $\phi:Gal(E \cdot E'/E) \to Gal(E'/E \cap E')$ is a well defined isomorphism. 
	\item The map $$\psi:Gal(E \cdot E' /K) \to Gal(E/K) \times Gal(E'/K)$$ mapping $g$ to $(g|E, g|E)$ is a well defined injective homomorphism and if $E \cap E' = K$ then $\psi$ is also surjective.
\end{enumerate}

\end{prop}

\begin{proof} As $E$ and $E'$ are Galois, it is easy to see that $E$ and $E'$ are splitting fields of two seperable (but not necessarily irreducible) polynomials with cofficients in $K$. Then $E \cdot E'$ is a splitting field of the lowest common multiple, which is also seperable and has coefficients in $K$. Let $g \in Gal(E \cdot E' /E)$ then for $e' \in E'$ we have that $g(e') \in E'$ as $E'$ is normal over $K$. Now if $e \in E \cap E'$ then $e \in E$ and thus $g(e) = e$. This shows that $g$ restricts to an element in $Gal(E' / E \cap E')$, so the homomorphism $\phi$ is well defined. If $g(e') = e'$ for all $e' \in E$ then $g$ acts trivially on $E \cdot E'$ as is already acts trivially on $E$. Thus this homomorphism has trivial kernel. We now show that surjectivity of $\phi$ as follows

\begin{align*} (E')^{Im \phi} &= \{ x \in E \cdot E' ~|~ x \in (E')^{Im \phi} \} \\
&= \{x \in E \cdot E' ~|~ x \in E' \text{ and } g(x) = x \text{ for all } g \in Gal(E \cdot E'/E) \} \\
&= \{ x \in E \cdot E' ~|~ x \in (E\cdot E')^{Gal(E \cdot E'/E)} \} \cap E' \\
&= E \cap E'  \end{align*}

where in the last equality we used the Galois correspondence. Thus by the Galois correspondence we have that $Gal(E' / (E\cap E')) = Im \phi$, thus showing that $\phi$ is surjective.

Now to show (2): Clearly the kernel of this map is trivial. Now suppose that $K = E \cap E'$. Let $(\sigma, \sigma') \in Gal(E/K) \times Gal(E'/K)$. Using part (1), this means that we may find extensions $\tilde{sigma} \in Gal(E \cdot E' /E')$ and $\tilde{\sigma}' \in Gal(E \cdot E'/E)$ of $\sigma$ and $\sigma'$ respectively. Now we claim that $\psi((\tilde{\sigma} \circ \tilde{\sigma}')) = (\sigma, \sigma')$. This is because for $e \in E$ we have $$(\tilde{\sigma} \circ \tilde{\sigma}')|_E(e) = \tilde{\sigma}(\tilde{\sigma}e) = \tilde{\sigma}(e) = \sigma(e) $$ where we have used the fact that $\tilde{\sigma}'  \in Gal(E \cdot E'/E)$ thus fixed $e$. A similair calculation verifies the second component of this identity, thus completing the proof of surjectivity. \end{proof}

\section{Cyclotomic fields}

We say that $\zeta \in \bC$ is a root of unity if $\zeta^n = 1$ for some $n>0$. We say that $\zeta$ is a primitive $n$-th root of unity if $\zeta^n = 1$ but $\zeta^m \neq 1$ for all $0<m<n$. In this section, we wish to understand cylcomotomic fields, i.e., field of the form $\bQ[\zeta_n]$ where $\zeta$.

\begin{lemma} The field $\bQ[\zeta_n]$ is a finite Galois extension of $\bQ$. The natural homomorphism $$Gal(\bQ[\zeta_n]/ \bQ) \to Aut(U_n) \cong (\bZ/n\bZ)^*$$ given by restriction to $U_n$ is injective, thus $$|Gal(\bQ[\zeta_n]/\bQ)| = |\bQ[\zeta_n]:\bQ|$$ divides $\phi(n)$. 

\end{lemma}

\begin{proof} It is a Galois extension as it is the splitting field of $X^n - 1$. Note that any $\bQ$-automorphism must permute the roots of unity. Moreover, this permutation induces an isomorphism of the multiplicative group $U_n \cong \bZ/n\bZ$. The endomorphisms of $\bZ/n\bZ$ are all of the form $x \mapsto ax$, and these are isomorphisms if and only if $gcd(a,n) = 1$. \end{proof}

We now strengthen the Lemma by showing that in fact $|\bQ[\zeta_n]:\bQ| = \phi(n)$.

\begin{thm} The field $\bQ[\zeta_n]$ is a finite Galois extension of $\bQ$. The natural homomorphism $$Gal(\bQ[\zeta_n]/ \bQ) \to Aut(U_n) \cong (\bZ/n\bZ)^*$$ given by restriction to $U_n$ is an isomorphism, thus $$|Gal(\bQ[\zeta_n]/\bQ)| = |\bQ[\zeta_n]:\bQ| = \phi(n).$$

\end{thm}

\begin{proof} Let $\zeta_n$ be a primitive $n$-th root of unity. Let $f(x) \in \bQ[x]$ be the monic minimal polynomial of $\zeta_n$. We claim that any other primitive $n$-th root of unity is also a root of $f$. For this, it is enough to show that for primes $p$ not dividing $n$ we have that $\zeta_n^p$ is also a root of $f$ (as any other primitive root of unity is of the form $\zeta_n^m$ for some $gcd(m,n)=1$, thus can be obtained by successively raising to such prime powers). Fix such a prime $p$. Suppose for contradiction that $f(\zeta_n^p) \neq 0$. Now as $$X^n - 1 = f(X)h(X)$$ for some $h(X) \in \bQ[x]$ monic, we get that $f(X)$ and $h(X)$ is monic (as $f(X)$ is monic) and thus by the Gauss Lemma we have that $f(X), h(X) \in \bZ[X]$. As $f(\zeta_n^p) \neq 0$ we have that $h(\zeta_n^p) = 0$. Thus $h(X^p) = f(X)g(X)$ for some $g(X) \in \bQ[x]$, which by the same argument must also be monic in $\bZ[x]$. We reduce this equation modulo $p$ to obtain $$(\overline{h}(x))^p = \overline{h}(x^p) = \overline{f}(x)\overline{g}(x)$$ in $\bF_p[x]$. Thus $\overline{f}(x)$ and $\overline{h}(x)$ must share a zero in some algebraic closure of $\bF_p$. Thus $x^n-1 = \overline{h}\overline{f}$ must have multiple zeros thus is not seperable. This however is only possibly if $p$ divides $n$ (as such a zero $\alpha$ would have to vanish on the derivative, i.e.,  $n\alpha^{n-1} = 0$ which implies that $\alpha^{n-1} = 0$ if $p$ does not divice $n$, so $\alpha=0$, but $0$ is not a root) , a contradiction.

Thus we have shown that $f(x)$ has at least $\phi(n)$ roots, but as $deg f$ divices $\phi(n)$, we have that the degree is exactly $\phi(n)$. So the Galois group also has $\phi(n)$ elements thus the injective homomorphism is an isomorphism. \end{proof}

\begin{prop} If $gcd(n,m)=1$ then $$\bQ[\zeta_n, \zeta_m] = \bQ[\zeta_{nm}]$$ and $$\bQ[\zeta_n] \cap \bQ[\zeta_m] = \bQ.$$ and there is an isomorphism $$Gal(\bQ[\zeta_{nm}]/\bQ) \to Gal(\bQ[\zeta_n]/\bQ) \times Gal(\bQ[\zeta_m]/\bQ)$$ mapping $\sigma \in Gal(\bQ[\zeta_{nm}]/\bQ)$ to the pair pair $(\sigma|_{\bQ[\zeta_n]}, \sigma|_{\bQ[\zeta_m]})$.

\end{prop}

\begin{proof} A simple calculation shows that if $gcd(n,m)=1$, then $\zeta_n\zeta_m$ is a primitive $nm$-th root of unity, thus the first equality. Now let $L = \bQ[\zeta_n] \cap \bQ[\zeta_m]$. Observe that $$\phi(n)\phi(m) = \phi(nm) = |\bQ[\zeta_{nm}:\bQ| = |\bQ[\zeta_{nm}:\bQ[\zeta_n]| \phi(n)$$ and thus $\bQ[\zeta_{n}, \zeta_m]:\bQ[\zeta_n]| = \phi(m)$. This means that $\zeta_m$ has degree $\phi(m)$ over the field $\bQ[\zeta_n]$. Thus $\zeta_m$ has degree at least $\phi(m)$ over the smaller field $L$, i.e., $|\bQ[\zeta_m]:L| \geq \phi(m)$. However $$\phi(m) = |\bQ[\zeta_m]:\bQ| = |\bQ[\zeta_m]:L||L:\bQ| \geq \phi(m) |L:\bQ|$$ and thus $L = \bQ$. Finally, the last claim follows from Proposition ??.  \end{proof}

We let $\phi_n(x) \in \bQ[x]$ denote the minimal polynomial of $\zeta_n$ over $\bQ$. Note that in the proof of ??? we saw that $$\phi_n (x) = \prod_{\zeta} (x - \zeta)$$ is a seperable polynomial of degree $\phi(n)$ such that each of the $\phi(n)$ primitive roots of unity are roots, thus the the product runs over the $\phi(n)$ different primitive roots of unity.

\begin{mydef} We call $\phi(n)$ the $n$-th cyclotomic polynomial.

\end{mydef}

\begin{prop} The cyclotomic polynomial $\phi_n(x)$ is a monic polynomial in $\bZ[x]$. We have the identity $$x^n - 1 = \prod_{d|n} \phi_d(x).$$

\end{prop}

\begin{proof} As $\phi_n(x)$ divides $x^n - 1$ in $\bQ[x]$ and is monic, we have that $x^n-1 = \phi_n(x) h(x)$ for some $h(x) \in \bQ[x]$ also monic. The Gauss lemma now shows that these polynomials must have integer coefficients. Finally, the identity follows because each root of unity is a primitive root of unity of some unique divisor $d$ of $n$. \end{proof}

We can use this identity to compute cyclotomic polynomials recursively, starting with the base case $$\phi_p(x) = \frac{x^p - 1}{x-1} = 1 + x + \cdots + x^{p-1}$$ for primes $p$.

\section{Norm and Trace}

Let $L/K$ be a finite field extension. For $a \in L$ we can define $Tr_{L/K}(a) \in L$ to be the trace of the $K$-linear map $\phi_a:L \to L$ given by $\phi_a(x) = ax$. We define the \textit{norm} $$N_{L/K}(a) = \operatorname{det}(\phi(a))$$  to be the determinant of this $K$-linear map. 

\begin{prop} Let $K$ be a field and let $\alpha \in \overline{K}$ be algebraic over. Then the characteristic polynomial of the $K$-linear map $\phi_{\alpha}:K(\alpha) \to K(\alpha)$ is precisely the minimal polynomial of $\alpha$ over $K$.

\end{prop}

\begin{proof} We know that the degree of the minimal polynomial is $|K(\alpha):K|$. Note also that for $P \in K[x]$ we have that $P(\phi_{\alpha}):K(\alpha) \to K(\alpha)$ is the zero map if and only if $P(\alpha) = 0$, thus $\phi_{\alpha}$ has the same minimal polynomial as $\alpha$ over $K$. But $|K(\alpha):K|$ is the dimension of $K(\alpha)$ of $K$, thus the minimal polynomial coincides with the characteristic polynomial.\end{proof}

Thus if $P(x) = \prod_{i=1}^n (X - \alpha_i)$ is the minimal polynomial of $\alpha$, then $$Tr_{K(\alpha):K}(\alpha) = \sum_{i=1}^n \alpha_i$$ and $$N_{K(\alpha):K}(\alpha) = \prod_{i=1}^n \alpha_i.$$

In case $P(x)$ is not separable, the $\alpha_i$ repeat with some multiplicity $q$ and we may write $$P(x) = \prod_{\rho} (X-\rho(\alpha))^q$$ where the product is over all $\rho \in Hom_{K}(K(\alpha): \overline{K})$. Note that $$q = |K(\alpha):K||K(\alpha):K|_s^{-1}.$$ Thus in this case, $$Tr_{K(\alpha)/K} = q \sum_{\rho} \rho(\alpha)$$ and $$N_{K(\alpha)/K}(\alpha) = \left(\prod_{\rho} \rho(\alpha) \right)^q.$$


\begin{prop} If $L/K$ is a finite extension and $\alpha$ in $K$, then $$Tr_{L/K}(\alpha) = |L:K(\alpha)| Tr_{K(\alpha)/K} (\alpha)$$ and $$N_{L/K} = \left( N_{K(\alpha)/K}(\alpha) \right)^{|L:K(\alpha)|}.$$

\end{prop}

\begin{proof} Let $y_1, \ldots, y_s \in L$ be a basis for $L$ over $K(\alpha)$, where $s = |L:K(\alpha)|$. Observe that $$L = \bigoplus_{i=1}^S K(\alpha) y_i$$ splits as a direct sum of $K$-vector spaces of dimension $|K(\alpha):K|$. Writing $\phi_{\alpha}:K(\alpha) \to K(\alpha)$ and $\psi_{\alpha}:L \to L$ to be the multiplication by $\alpha$ maps (both viewed as $K$-linear maps on $K$-vector spaces), we see that for by writing each $x \in L$ as $x = (x_1, \ldots, x_s)$ with respect to this decomposition we have $$\psi_{\alpha} x = (\phi_{\alpha} x_1, \ldots, \phi_{\alpha} x_s)$$ and thus $$Tr (\psi_{\alpha}) = s \cdot Tr (\phi_{\alpha})$$ and $$\operatorname{det} (\psi_{\alpha}) = \operatorname{det} (\phi_{\alpha})^s,$$ as required. \end{proof}

\begin{thm} Let $L/K$ be a finite extension and let $\alpha \in K$. Let $r=|L:K|_s$ and let $\sigma_1, \ldots, \sigma_r$ be the distinct elements of $Hom_K(L, \overline{K})$, i.e., the homomorphisms $L \to \overline{K}$ that fix $K$. Then $$Tr_{L/K}(\alpha) = \frac{|L:K|}{|L:K|_s} \sum_{i=1}^r \sigma_i (\alpha)$$ and $$N_{L/K}(\alpha) = \left(\prod_{i=1}^r \sigma_i (\alpha) \right ) ^{\frac{|L:K|}{|L:K|_s}}.$$

\end{thm}

\begin{proof} For each $\rho \in Hom_{K}(K(\alpha), \overline{K})$, fix an extension $\overline{\rho}:\overline{K} \to \overline{K}$ of $\rho$. Note that each $\sigma \in Hom_{K}(L, \overline{K})$ may be uniquely written by (the proof of) Proposition CITE in the form $\overline{\rho} \circ \tau$ where $\tau \in Hom_{K(\alpha)}(L : \overline{K})$. As such a $\tau$ must fix $\alpha$ we have $(\overline{\rho} \circ \tau) (\alpha) = \rho(\alpha)$ and so we can write 

\begin{align*}Tr_{L/K}(\alpha) &= |L:K(\alpha)| \frac{|K(\alpha):K|}{|K(\alpha):K|_s} \sum_{\rho} \rho(\alpha) \\ 
&= |L:K(\alpha)| \frac{|K(\alpha):K|}{|K(\alpha):K|_s} \frac{1}{|L:K(\alpha)|_s} \sum_{\tau} \sum_{\rho} (\overline{\rho} \circ \tau)(\alpha) \\
&= \frac{|L:K|}{|L:K|_s} \sum_{\sigma} \sigma(\alpha), \end{align*} where a summation over $\sigma$ is over $\sigma \in  Hom_{K}(L, \overline{K})$, a summation over $\tau$ is over $ \tau \in Hom_{K(\alpha)}(L : \overline{K})$ and a summation over $\rho$ is over  $\rho \in Hom_{K}(K(\alpha), \overline{K})$.

The proof for the norm is similair.\end{proof}

Note that as a corollary, we get that if $L/K$ is a finite Galois extension then $$Tr_{L/K}(\alpha) = \sum_{\sigma \in Gal(L/K)} \sigma(\alpha)$$ and  $$N_{L/K}(\alpha) = \prod_{\sigma \in Gal(L/K)} \sigma(\alpha).$$
\begin{corol}\label{corol: galois fixed trace and norm} Let $L/K$ be a finite Galois extension, then $$Tr_{L/K} \circ \sigma = Tr{L/K}$$ and $$N_{L/K} \circ \sigma = N_{L/K}.$$

\end{corol}

If $L/K$ is a finite extension, we can define a bilinear form $tr:L \times L \to K$ given by $tr(x,y) = Tr_{L/K}(xy)$. Note that if $L/K$ is finite Galois, then the previous result says that the Galois group preserves this bilinear form.

\begin{prop} Let $L/K$ be a finite extension. Then $L/K$ is seperable if and only if $Tr_{L/K}:L \to K$ is a non-trivial (hence surjective) linear functional. If $L/K$ is seperable, then the bilinear form $tr$ is non-degenerate, i.e., if $x \in L$ is such that $tr(x,y) = 0$ for all $y \in Y$, then $y=0$.

\end{prop}

\begin{proof} If $L/K$ is not seperable, then $q = |L:K||L:K|_s^{-1}$ must be a power of $char K \neq 0$ (see Proposition~\ref{prop: seperable degree prime power} and Theorem~\ref{thm: characterise seperable}). Thus we get that $Tr_{L/K} = q \sum_{\sigma} \sigma$ is identically zero. Now assume $L/K$ is seperable, thus $q = 1$. Now the elements $\sigma \in Hom_K(L, \overline{K})$ are distinct characters (multiplicative maps) $L^* \to \overline{K}^*$, thus linearly independent and so $Tr_{L/K}$ cannot be identitically zero. It now follows immediately that the bilinear form is non-degenerate, for if $x \in L$ with $x \neq 0$ and $tr(xy) = 0$ for all $y \in L$, then as $y$ is a field we have $xL=L$ and so $Tr_{L/K}$ vanishes on $L$, but we have just shown that it does not. \end{proof}

\section{Cyclic Galois Extensions}

\begin{thm}[Hilbert 90] Let $L/K$ be a finite cyclic Galois extension and let $\sigma \in Gal(L/K)$ be a generator. For $b \in L$, we have that $N_{L/K}(b) = 1$ if and only if there exists non-zero $a \in L$ such that $$b = a (\sigma(a))^{-1}.$$

\begin{proof} If $b = a (\sigma(a))^{-1}$ for some non-zero $a \in L$ then by multiplicativity of $N_{L/K}$ and invariance under $\sigma$ (Corollary~\ref{corol: galois fixed trace and norm}) we get that $$N_{L/K}(b) = N_{L/K}(a) \left( N_{L/K}(\sigma(a)) \right)^{-1} = N_{L/K}(a) (N_{L/K}(a))^{-1} = 1.$$

Conversely, suppose now that $N_{L/K}(b)=1$. Let $$u_j = \prod_{j=0}^i \sigma^{i}(b).$$ We note that $u_j$ is $n$-periodic, where $n = Gal(L/K)$ as $$1 = N_{L/K}(b) = \prod_{\sigma \in Gal(L/K)} \sigma(b) = u_{i+n}u_i^{-1}$$ and furthermore all $u_j$ have norm $1$ thus are non-zero. It follows, from the linear independence of the maps in $Gal(L/K)$, that $$\sum_{i=0}^{n-1} u_j \sigma^{j}$$ is not the zero function. Thus there exists a $c \in L$ for which $$a := \sum_{i=0}^{n-1} u_j \sigma^{j}(c) \neq 0. $$ Now we have 

\begin{align*} b \sigma(a) &= \sum_{i=0}^{n-1} b\sigma(u_i) \sigma^{i+1}(c) \\
&= \sum_{i=0}^{n-1} u_{i+1} \sigma^{i+1}(c) \\
&= \sum_{i=1}^n u_{i} \sigma^{i}(c)\\
&= a \end{align*}

where in the final equality we used the periodicity of the summands, $\sigma^{i+n} = \sigma^i$ and $u_{i+n} = u_i$. \end{proof}

\end{thm}

\begin{eg} Consider the degree two extension $\bC/\bR$. It is Galois with Galois group generated by $\sigma(z) = \overline{z}$ being complex conjugation. We have $$N_{\bC/\bR}(b) = b \sigma(b) = |b|^2.$$

On the other hand we have $$a \sigma(a)^{-1} = \frac{a^2}{\overline{a}{a}} = \frac{a^2}{|a^2|},$$ which has norm $1$. Thus $b$ takes the form $a \sigma(a)^{-1}$ if and only if it has norm $1$.

Now consider instead the extension $\bQ[i]/\bQ$. It is also Galois of degree $2$ generated by $\sigma:\bQ[i] \to \bQ[i]$ being complex conjugation. Again $$N_{\bQ[i]/\bQ[]} (x+iy) = |x+iy|^2 = x^2 + y^2.$$ Thus Hilbert 90 shows gives an explicit paramatrization of the rational solutions $x+iy \in \bQ[i]$ to $x^2 + y^2 = 1$, they are all of the form $$\frac{u+iv}{u-iv} \quad u, v \in \bQ.$$\end{eg}

This theorem allows us to relate cyclic Galois extensions with radical extensions (adjoining an $nth$-root).

\begin{prop} Let $L/K$ be a cyclic Galois extension of degree $n$ such that $K$ contains a primitive $n$-th root of unity. Then $L = K(a)$ for some non-zero $a \in L$ with $a^n \in K$ and minimal polynomial $X^n - a^n \in K[x]$.

\end{prop}

\begin{proof} Let $\zeta$ be a primitive $n$-th root of unity. Note that $N_{L/K}(\zeta) = \zeta^n = 1$ and thus by Hilbert 90 we have a non-zero $a \in L$ such that $\sigma(a) = \zeta a$, where $\sigma \in Gal_{L/K}$ is a generator. Now observe that $\sigma(a^n) = \sigma(a)^n = \zeta^n a^n = a^n$. Thus $a^n$ is fixed by $\sigma$, which means that $a^n \in K$. Thus $a$ is a root of $g(x) = x^n - a^n \in K[x]$. Observe that $a_i := \sigma^i(a) = \zeta^i a$ and thus these elements are distinct as $\zeta$ is primitive. If $f(x) \in K[x]$ is a minimal polynomial for $a$, then $f(\sigma^i(a)) = f(a) = 0$ and thus $f(x)$ has at least $n$ distinct roots, thus is of degree at least $n$. As $f(x) | g(x)$ we have that $g(x) = f(x)$ and so $x^n - a^n \in K[x]$ is indeed the minimal polynomial of $a$ over $K$. This also means that $|K(a):K| = n = |L:K|$, thus $L = K(a)$. \end{proof}

\begin{eg} Let us show that the assumption that $K$ contains a primitive $n$-th root of unity cannot be dropped. Let $L = \bQ(x_1,x_2,x_3)$ and let $\sigma(x_i) = x_{i+1}$ be a cyclic permutation of the variables and let $K = L^{G}$ where $G$ is the cyclic group generated by $\sigma$. Then we know that $L/K$ is a Galois extension with Galois group $G$ (use Proposition~\ref{prop: fixed field galois group} as $G$ is finite). Now we show that $L$ cannot have a primitive $n$-th root of unity. The easiest way of showing this is to embedd $L$ into $\bR$ via a field homomorphism (choose three algebraically independent numbers). Thus if $a^3 = 1$ in $L$ then $a=1$. Now if $L=K(a)$ for some $a \in L$ with $a^3 \in K$ then $(\sigma(a))^3 = a^3$ and so $a = \sigma(a)$, thus $a \in K$ which means that $K = L$, a contradiction. \end{eg}

\begin{prop}\label{prop: radicals implies cyclic galois} Let $K$ be a field with a primitive $n$-th root of unity and suppose that $L$ is a field with $L=K(a)$ for some $a \in L$ such that $a^n \in K$. Then $L/K$ is a finite cyclic Galois extension with degree $d = [L:K]$ where $d | n$ and $a$ has minimal polynomial $x^d - a^d \in K[x]$.

\end{prop}

\begin{proof} We assume $a \neq 0$, otherwise the result is obvious. Now let $\zeta \in K$ be a primitive $n$-th root of unity and let $a_j = \zeta^j a$. Observe that $a_0, \ldots, a_{n-1}$ must be distinct ($a \neq 0$ and $\zeta$ is a primitive $n$-th root) and satisfy that $a_j^n - a^n =0$. Thus $X^n - a^n$ has $n$ distinct roots in $L$, which shows that $L=K(a)$ is seperable over $K$, thus $L/K$ is Galois. Now the Galois group permutes that $a_j$, thus for $\sigma \in Gal_{L/K}$ we have $$\phi(\sigma) := \frac{\sigma(a)}{a} \in U \subset K^{*}$$ where $U$ is the group of order $n$ generated by $\zeta$. We claim that $\phi:Gal(L/K) \to U$ is a group homomorphism. This is easy to see as $$\phi(\sigma_1) \phi(\sigma_2) = a^{-1} \sigma_1(a) \phi(\sigma_2) = a^{-1}\cdot \sigma_1(a \cdot \phi(\sigma_2)) = a^{-1}\sigma_1(\sigma_2(a)).$$ Moreover, this mapping is injective as $L=K(a)$ so any $K$-homomorphism is determined by $a$. Thus $Gal(L/K)$ is isomorphic to a subgroup of $U$, thus is cyclic of degree $d | n$. Let $\sigma$ be a generator, thus $\sigma$ permutes those elements $a_j$ where $j$ is divisibly by $r = n/d$, therefore $$g(x) = \prod_{j=0}^{d-1}(X - \sigma^j(a)) = \prod_{j=0}^{d-1}(X - \zeta^{rj} a)$$ must be an element of $K[x]$ (as it is fixed by $\sigma$) thus it is the minimal polynomial of $a$, as it has degree $d = Gal(L/K) = |L:K|$. It now remains to show that $g(x) = x^d - a^d$. This must be the case since the right hand side has at most $d$ roots while $g(x)$ has exactly $d$ distinct roots of $x^d - a^d$. \end{proof}

\begin{eg} If $a$ is a primitive $12$-th root of unity in $\bC$, that $L = \bQ(a)$ is the splitting field of $X^{12} - a$. Thus $L$ is Galois over $\bQ$. However $Gal(L/\bQ)$ is not cyclic, it is isomorphic to $(\bZ/4)^* \times (\bZ/3)^* = \bZ/2 \times \bZ/2$. This does not contradict the theorem, as $\bQ$ does not have a primitive $12$-th root of unity.\end{eg}

\begin{eg} $\bQ(2^{1/4})$ not even Galois over $\bQ$, again, not a contradiction as $\bQ$ does not have a primite $4$-th root of unity. On the other hand $L = \bQ[i](2^{1/4})$ is a Galois extension of $K = \bQ[i]$ and the Galois group is cyclic of order $4$. 

\end{eg}

\section{Solvable groups}

\begin{mydef} Let $G$ be a group. Define the commutator $[a,b] = aba^{-1}b^{-1}$. For subgroups $H, H' \leq G$ define the commutator subgroup $[H,H']$ to be the group generatated by $\{[h,h'] ~|~ h \in H h' \in H\}$.

\end{mydef}

\begin{prop} Let $G$ be a group. Then $[G,G]$ is a normal subgroup of $G$ and $G/[G,G]$ is abelian. Moreover, if $H \leq G$ is a normal subgroup of $G$ such that $G/H$ is abelian, then $[G,G] \leq H$. Thus $[G,G]$ is the minimal normal subgroup $H \leq G$ such that $G/H$ is abelian.

\end{prop}

\begin{proof} Note that $$g[a,b]g^{-1} = gaba^{-1}b^{-1}g^{-1} =  gag^{-1}gbg^{-1}ga^{-1}g^{-1}gb^{-1}g^{-1} = [gag^{-1}, gbg^{-1}. $$ Thus the generating set is conjugation invariant. Also, note that $$[a,b]^{-1} = (aba^{-1}b^{-1})^{-1} = bab^{-1}a^{-1} = [b,a]^{-1}$$ thus any element in $[G,G]$ can be written as a product of commutators (no need to consider inverses of commutators). Thus since the commutators are conjugation invariant, $[G,G]$ is normal. Now let $\pi:G \to G/[G,G]$ be the quotient map. Then $$[\pi(g), \pi(h)] = \pi(g)\pi(h)\pi(g)^{-1}\pi(h)^{-1} = \pi([g,h]) = 1$$ thus all elements of this quotient group commute. Finally, suppose that $H$ is a normal subgroup of $G$ with $G/H$ abelian. If $\pi:G \to G/H$ then $1 = [\pi(g), \pi(h)] = \pi([g,h])$ and so $[g,h] \in ker~\pi = H$. Thus $[G,G] \subset H$. \end{proof}

\begin{mydef} Define $D^{0}G =G, DG = [G,G]$ and $D^{i+1}G = D (D^{i}G) = [D^iG, D^iG]$. Thus we have a series $$G \geq D^1 G \geq D^2 G \geq \ldots$$ where the successive quotients are normal. We say that $G$ is \textit{solvable} if $D^nG = \{1\}$ for some $n \geq 0$.

\end{mydef}

\begin{prop} A group $G$ is solvable if and only if there is a finite sequence $$G_0 = G \geq G_1 \geq G_2 \ldots \geq G_n = \{1\}$$ such that $G_{i+1}$ is normal in $G_i$ and $G_i/G_{i+1}$ is abelian.

\end{prop}

\begin{proof} Given such a normal series with abelian quotients, we can show that $D^i G \leq G_i$ by induction as follows. Clearly true for $i=0$ as we have equality. Suppose now that $D^i G \leq G_i$ for some $i < n$. Then $D^{i+1}G = D(D^iG) \leq DG_{i} = [G_{i}, G_{i}]$. But since $G_i /G_{i+1}$ is abelian, we have by the previous Proposition that $[G_{i}, G_{i}] \leq G_{i+1}$. Thus $D^{i+1}G \leq G_{i+1}$. The converse is clear.\end{proof}

\begin{eg} The symmetric group $S_5$ is not solvable. We will show this by showing that $[S_5, S_5] = A_5$ and $[A_5, A_5] = A_5$, thus the series will never reach the trivial subgroup. Let $sign:S_5 \to \{-1,1\}$ denote the sign map. It is clear that $[S_5, S_5] \leq A_5$ as $$sign([g,h]) = sign(g)sign(h)sign(g)^{-1}sign(h)^{-1} = 1.$$ For the reverse inclusion, we note that $A_5$ can be generated by $3$-cycles as for distinct $a,b,c,d \in \{1, \ldots, 5\}$ we have $$(a~b)(b~c) = (a~b~c)$$ and $$(a~b)(c~d) = (a~b~c)(b~c~d)$$, thus it is enough to show that any $3$-cycle can be expressed as a commutator. To see this, note that $$(a~b~c) = (a~c)(b~c)(a~c)^{-1}(b~c)^{-1}.$$ This shows that $[S_5,S_5] = A_5$. We now show that $[A_5, A_5] = A_5$. Thus it will be enough to express a $3$ cycle as a commutator of $3$ cycles. This is because $$(1~2~3) = (1~2~4)(1~3~5)(1~2~4)^{-1}(1~3~5)^{-1}.$$

Thus $S_5$ is not solvable.
\end{eg}

\begin{prop} Let $H \leq G$. Then if $G$ is solvable then so is $H$. If $H$ is normal in $G$, then $G$ is solvable if and only if $H$ and $G/H$ are solvable.

\end{prop}

\begin{proof} First part is obvious as $D^iH \leq D^i G$. Now suppose that $H$ is normal in $G$. Then the surjective map $\pi:G \to G/H$ satisfies that $D^i(\pi(G)) =\pi(D^i(G))$, thus if $G$ is solvable then so is $H$. Now if $G/H$ is solvable, then $$\pi(D^n(G)) = D^n(\pi(G)) = \{1\}$$ and thus $D^n(G) \leq H$ for some $n \geq 0$. Now if $H$ is solvable, then $$D^{n+m}(G) =D^m(D^n(G)) \leq D^mH = \{1\}$$ for some $m \geq 0$ and thus $G$ is solvable. \end{proof}

\section{Solvability by radicals}

\begin{mydef} Let $L/K$ be a field extension in characteristic $0$. We say that $L/K$ is \textit{solvable by radicals} if $L \leq E$ for some field $E$ such that there is a sequence $$E_0 = K \leq E_{1} \leq E_2 \ldots \leq E_n = E$$ of field extensions where $E_{i+1} = E_i(a)$ for some $a \in E_{i+1}$ such that $a^n \in E_i$ for some $n \geq 1$.

\end{mydef}

\begin{mydef} Let $L/K$ be a finite field extension. Then we say that $L/K$ is solvable if there is a finite extension $L \leq \tilde{L}$ such that $\tilde{L}/K$ is Galois with solvable Galois group.

\end{mydef}

\begin{proof} If $L/K$ is a finite Galois extension that is solvable. Then $Gal(L/K)$ is solvable.

\end{proof}

\begin{proof} As $L/K$ is solvable, there is a Galois extension $\tilde{L}/K$ with $K \leq L \leq \tilde{L}$ such that $G = Gal(\tilde{L}/K)$ is solvable. But as $L/K$ is assumed to be Galois, we know that $Gal(L/K)$ is a quotient of $G$ and thus also solvable. \end{proof}

\begin{lemma} Let $L/K$ be a finite field extension in characteristic $0$. Suppose that $K \leq F$ is a field extension and that $L \leq \overline{F}$ (the algebraic closure of $F$). 

\begin{enumerate}
	\item If $L/K$ is solvable, then $FL/F$ is solvable.
	\item If $L/K$ is solvable by radicals, then $FL/F$ is solvable by radicals.
\end{enumerate}

\end{lemma}

\begin{proof} Proof of (1): As $L/K$ is solvable, there is a field extension $L \leq \tilde{L} \leq \overline{F}$ such that $K \leq \tilde{L}$ is finite Galois with $Gal(\tilde{K}/K)$ solvable. Now $\tilde{L} = K[A]$ where $A$ is the set of roots of some finite set of polynomials of $K[x]$ (by definition of normal). Now $F\tilde{L} = FK[A] = F[A]$, thus $F\tilde{L}$ is Galois over $F$ and $F\tilde{L} \geq F \geq K$. Thus it is enough to show that $Gal(F\tilde{L}/F)$ is solvable. Note that there is a restriction homomorphism $$Gal(F\tilde{L}/F) \to Gal(\tilde{L}/K)$$ which is well defined as $\tilde{L}/K$ is normal and any $\sigma$ that fixes $F$ must fix also $K \leq F$. Note also that this map is injective since if $\sigma \in Gal(F\tilde{L}/F)$ acts trivially on $\tilde{L}$ then it also acts trivially on $F\tilde{L}$ as it already fixed $F$. Thus $Gal(F\tilde{L}/F)$ is a subgroup of a solvable group, thus solvable.

Now to prove (2): If $L/K$ is solvable by radicals, then there is a field $L \leq E \leq \overline{F}$ such that $E = K(a_1, \ldots, a_n)$ where $a_i^{n_i} \in K(a_1, \ldots, a_{i-1})$ for some $n_i \geq 1$. It follows that $F \leq FL \leq FE$ is solvable by radicals since $$FE = FK(a_1, \ldots, a_n) = F(a_1, \ldots a_n)$$ with $a_i^{n_i} \in K(a_1, \ldots, a_{i-1}) \subset F(a_1, \ldots, a_{i-1}).$ \end{proof}

\begin{prop} Let $K \leq L \leq M$ be finite field extensions in characteristic $0$. Then $K \leq M$ is solvable if and only if $K \leq L$ and $L \leq M$ are solvable.

\end{prop}

\begin{proof} We first prove (1): Suppose $K \leq M$ is solvable. Thus there is an $M' \geq M$ finite set so that $K \leq M'$ is galois and $Gal(K/M')$ is solvable. As $M'$ is an extension of $L$, this means that $L/K$ is also solvable. Finally $M'/L$ is also Galois with $Gal(M'/L)$ a subgroup of $Gal(M/K)$, thus solvable. Therefore $M/L$ is solvable.

Conversely, suppose that $K \leq L$ and $L \leq M$ is solvable. We first show that we may reduce to the case where $K \leq L$ and $L \leq L$ are also Galois with solvable Galois group: We know there is a finite extension $L' \geq L$ such that $L'/K$ is Galois and solvable. We now apply the previous lemma with our $L$ playing the role of $K$ (in the theorem), our $L'$ playing the role of $F$ and $M$ playing the role of $L$. Thus by assumption $M/L$ is solvable, so the theorem applies to show that $ML'/L'$ is solvable. By definition, this means we have an $M' \geq ML'$ such that $M'/L'$ is Galois with solvable Galois group. Thus we have a chain $K \leq L' \leq M'$ of Galois extensions with Galois groups. Thus if we can show that $M'/K$ is solvable, then $M/K$ is solvable as desired.

Thus we have reduced to the case that $K \leq L$ is Galois with solvable Galois group and $L \leq M$ is Galois with solvable Galois group and we wish to show that $K \leq M$ is solvable (not necessarily Galois). Now let $M \leq M'$ be the normal closure of $M/K$. By construction, recall the construction of $M'$ as follows: Write $M = K(a_1, \ldots, a_q)$ for some $a_i \in M$. Now $M' = K[\mathcal{A}]$ where $$\mathcal{A} = \{ a \in \overline{K} ~|~ a \text{ is the root of the minimal polynomial of some } a_i \}.$$ Thus $M'/K$ is finite and $$M' = \prod_{\sigma \in Gal(M'/K)} \sigma(M)$$ is the field generated by all the $\sigma(M)$ as $\sigma$ ranges in $Gal(M'/K)$. As $$Gal(L/K) \cong Gal(M'/K) / Gal(M'/L)$$ it suffices to show that $Gal(M'/L)$ is solvable as $Gal(L/K)$ is solvable. Now define a map $$Gal(M'/L) \to \prod_{\sigma \in Gal(M'/K)} Gal(\sigma(M)/L)$$ given by restricting $g \in Gal(M'/L)$ to the tuple of restrictions $g|_{\sigma(M)}$. Let us show that this is well defined. We have to show that $g \sigma(M) = \sigma(M)$. Note that $gM =M$ as $M/L$ is normal. Now since $Gal(M'/L)$ is normal we have that $\sigma^{-1} g \sigma \in Gal(M'/L)$ and so again $ \sigma^{-1} g \sigma M = M$, thus $g \sigma M = \sigma M$ as desired. This shows that this map is well defined. Now we show that it is injective. Thus suppose that $g$ fixes each element of each $\sigma M$. But then $g$ fixes each element of $M'$ as these fields $\sigma M$ generated $M'$. Thus $Gal(M'/L)$ is a subset of this product of Galois groups, it thus suffices to show that each $Gal(\sigma M/L)$ is solvable (because a finite product of solvable groups is solvable). We know $Gal(M/L)$ is solvable and we claim $Gal(\sigma(M)/L)$ is isomorphic to it via the isomorphism $g \mapsto \sigma g \sigma^{-1}$. To see that it is well defined, note that $\sigma g \sigma^{-1} \sigma M = \sigma gM = \sigma M$ and that if $\ell \in L$, then $\sigma^{-1} \ell \in L$ since $L/K$ is normal and thus $g$ fixes $\sigma^{-1}\ell$, which gives $$\sigma g \sigma^{-1} \ell = \sigma \sigma^{-1} \ell = \ell$$, thus $\sigma g \sigma^{-1}$ does indeed fix $L$. Thus shows that this map is well defined, and it is easy to see that it is an isomorphism. \end{proof}

\begin{thm} If $K \leq L$ is a finite field extension in characteristic $0$ that is solvable by radicals, then $K \leq L$ is solvable (i.e., $Gal(L',K)$ is solvable for some $L' \geq L$ with $K \leq L'$ a finite galois extension).

\end{thm}

\begin{proof} By the previous proposition, it suffices to show that if $K(a)/K$ is solvable where $a \in L$ is such that $a^n \in K$ for some $n$. Now consider the field extension $K \leq K(\zeta) \leq K(\zeta)(a)$ where $\zeta$ is a primitive $n$-th root of unity. Thus $K(\zeta)/K$ is solvable (it is a Galois extension with abelian galois group) while $K(\zeta)(a)/K(\zeta)$ is a finite Galois extension with cylic Galois group by Proposition~\ref{prop: radicals implies cyclic galois}. Thus by the previous proposition, $K \leq K(\zeta)(a)$ is solvable, thus $K(a)/K$ is solvable.

\end{proof}

\begin{thm} There is no quintic formula. More precisely, let $L = F[x_1, \ldots, x_5]$, where $F$ is a field of characteristic zero, and let $K = L^{S_5}$ where $S_5$ acts on $\{x_1, \ldots, x_5\}$ be permutatations. Then the extension $L/K$ is not solvable by radicals.

\end{thm}

\begin{proof} The Galois Group of $L/K$ is $S_5$, hence not solvable.\end{proof}



\end{document}